Palaeodistribution of crop progenitors and wild foods in the Southern Levant

### 2023-08-03 analysis/analysis.Rmd
Switch to stars from raster/terra

```{r dependencies}
library("dplyr")
library("ggplot2")
library("ggspatial")
library("here")
library("parsnip")
library("purrr")
library("ranger")
library("rpaleoclim")
library("sf")
library("tibble")
library("tidyr")

@@

```{r data-paleoclim}
tibble(period = c("lgm", "hs1", "ba", "yds", "eh", "cur")) |>
  mutate(raster = map(period, paleoclim, "2_5m", 
                      region = buffer_bbox(w_asia, 10),
                      cache_path = here(".paleoclim_cache"))) -> # TODO: revert to default cache
  climate

training_climate <- filter(climate, period == "cur")$raster[[1]]

occ |>
  mutate(bio = terra::extract(training_climate, data.frame(decimalLongitude, decimalLatitude), 
                              method = "bilinear", ID = FALSE)) |>
  unpack(bio) |>
  ungroup() ->
  occ
```

@@

```{r model}
rand_forest(mode = "regression") |>
  fit(as.numeric(present) ~ ., data = select(occ, present, starts_with("bio_"))) ->
  rf_fit
```

@@

```{r predict}
predict(rf_fit, as.data.frame(filter(climate, period == "cur")$raster), type = "raw")
```

>>>>

```{r dependencies}
library("dplyr")
library("fs")
library("ggplot2")
library("ggspatial")
library("here")
library("parsnip")
library("purrr")
library("ranger")
library("rpaleoclim")
library("sf")
library("stars")
library("stringr")
library("tibble")
library("tidyr")

@@

```{r data-climate}
raw_data("paleoclim") |>
  dir_ls() |>
  map(dir_ls, glob = "*.tif") |>
  map(read_stars) |>
  map(\(x) set_names(x, path_ext_remove(path_file(names(x))))) ->
  climate

# Tidy up names
names(climate) |>
  basename() |>
  str_remove(regex("chelsa_", ignore_case = TRUE)) |>
  str_split_i("_", 1) ->
  names(climate)
```

```{r training-climate}
occ |>
  mutate(bio = extract_points(climate$cur, occ$geometry)) |>
  unpack(bio) ->
  occ
```

@@

```{r model}
# TODO: Investigate where these NAs are coming from
rand_forest(mode = "regression") |>
  fit(as.numeric(present) ~ ., data = select(occ, present, starts_with("bio_"))) ->
  rf_fit
```

@@

```{r predict}
# TODO: deal with missing data in pred vars
# TODO: turn back to something spatial? stars?
predict(rf_fit, climate$cur)
```
